Privacy-Respecting Type Error Telemetry at Scale

### 2023-12-06 programming23/paper.tex
artifact todo outline, edit buckets/hours discussion

\Cref{f:records-per-hour} provides a time-ordered distribution of the data.
The x-axis is labeled with a California time zone; the thin vertical lines
mark the start of a new day.
Each day has 24 records, one per hour, represented by a blue bar
indicating the number of records generated in that hour.
Each hour has a few dozen to over 3,000 records, typically about
1,000.\footnote{We conjecture the very low bars are due to outage on
  the logging server, but need to investigate further.}
%% \FILL{} investigate one tower, what's happening?? all one session?
Telemetry ended in early April; the long tail is due to developers
updating Roblox Studio later.
% BEN: I see no basis for this; nor is this at all interesting.
%% ben: basis = it's where the peaks are
% We do not know which timezone the records originated in, but since the counts
% tend to peak midday California time it seems that the majority
% of Roblox developers are following a Western Hemisphere schedule.
The regular pattern of the peaks suggests that many creators
follow a regular schedule; these may be
professional developers.
The tall weekend (shaded) regions could be because
Roblox has a significant school-aged creator community.

@@

This paper presents the first large-scale analysis of developers' interactions with
an industrial strength typechecker. It is nevertheless intentionally
limited by the nature and frequency of sampling. Despite these restrictions,
the analysis offers several lessons for the Luau
developers and other maintainers of gradually-typed languages:
\begin{itemize}
  \item
    Nearly all sessions (\pct{90}) use the default ``no check''
    typechecking mode~(\cref{f:dataset-overview}).
    Since the next typing mode (\mnonstrict{}) requires no annotations and is
    very conservative about the errors it reports, we conjecture that
    developers would have no objections to using it and are simply unaware that
    other modes exist.
%  \item
%    \pct{50} of all errors are in current module, so probably errors appear locally (not in a distant module);
%    \pct{5} of (current? old??) type errors overlap with edit range, so they don't stick around.
  \item
    Sessions tend to fix errors immediately rather than dealing with them in a batch
    or ignoring the errors altogether~(\cref{f:error-density,f:error-changes}).
    This should give Roblox confidence to make \mnonstrict{} the default in the future.
    % Changes to the number of errors balance one another~(\cref{f:error-density});
    % edits are equally likely to add or remove errors~(\cref{f:error-changes}).
  \item
    All Luau scripts interact with data assets, but no typechecking mode
    has a precise view of the data.
    Improving precision is an important direction for future work.
    Until then, \mstrict{} mode clearly needs to relax its data model
    analysis because it raises even more errors than \FS{} checking~(\cref{s:tefs-compass}).
\end{itemize}

>>>>

\Cref{f:records-per-hour} provides a time-ordered distribution of the data.
The x-axis is labeled with a California time zone; the thin vertical lines
mark the start of a new day.
For each of the 24 hours in each day, there is one blue bar
indicating the number of records generated during that hour.
Each hour has a few dozen to over 3,000 records, typically about
1,000.\footnote{We conjecture the very low bars are due to outage on
  the logging server, but need to investigate further.}
%% \FILL{} investigate one tower, what's happening?? all one session?
Telemetry ended in early April; the long tail is due to developers
updating Roblox Studio later.
% BEN: I see no basis for this; nor is this at all interesting.
%% ben: basis = it's where the peaks are
% We do not know which timezone the records originated in, but since the counts
% tend to peak midday California time it seems that the majority
% of Roblox developers are following a Western Hemisphere schedule.
The regular pattern of the peaks suggests that many creators
follow a regular schedule; these may be
professional developers.
The tall weekend (shaded) regions could be because
Roblox has a significant school-aged creator community.

@@

This paper presents the first large-scale analysis of developers' interactions with
an industrial strength typechecker in a privacy-respecting way.
Despite the intentionally-restricted data,
the analysis offers several lessons for the Luau
developers and other maintainers of gradually-typed languages:
%% TODO more lessons?
\begin{itemize}
  \item
    Nearly all sessions (\pct{90}) use the default ``no check''
    typechecking mode~(\cref{f:dataset-overview}).
    Since the next typing mode (\mnonstrict{}) requires no annotations and is
    very conservative about the errors it reports, we conjecture that
    developers would have no objections to using it and are simply unaware that
    other modes exist.
%  \item
%    \pct{50} of all errors are in current module, so probably errors appear locally (not in a distant module);
%    \pct{5} of (current? old??) type errors overlap with edit range, so they don't stick around.
  \item
    Sessions tend to fix errors immediately rather than dealing with them in a batch
    or ignoring the errors altogether~(\cref{f:error-density,f:error-changes}).
    This should give Roblox confidence to make \mnonstrict{} the default in the future.
    % Changes to the number of errors balance one another~(\cref{f:error-density});
    % edits are equally likely to add or remove errors~(\cref{f:error-changes}).
  \item
    All Luau scripts interact with data assets, but no typechecking mode
    has a precise view of the data.
    Improving precision is an important direction for future work.
    Until then, \mstrict{} mode clearly needs to relax its data model
    analysis because it raises even more errors than \FS{} checking~(\cref{s:tefs-compass}).
\end{itemize}
